working title Compatibility system and stigma size are the main predictors of heterospecific pollen effect

### 2019-02-19 Manuscript_draft/Manuscript_draft.Rmd
some edits 3rd paragraph

**Paragraph 1** General idea to our concept

@@

**Paragraph 2** Introducing topic and knowledge gap

@@

**Paragraph 3** Expanding ideas with examples

Incompatibility system is another plant trait that appears to play an important role in foreign pollen effect. Species that are self incompatible have stronger barriers to heterospecific pollen than self-compatible species @ashman2013. However, note the large variability in mating systems across populations @whitehead2018 [...](). The type of incompatibility, (i.e. whether sporophytic or gametophytic) is related to the location of pollen recognition; sporophytic incompatibility relates to signaling at the stigma surface while gametophytic occurs within the style @barrett1988.  This later acting pollen recognition mechanism is associated with greater negative effect than sporophytic recognistion @ashman2013. [if you introduce this,I eexpect you will use it in your analysis, remove]() [Nonetheless, an effect of foreign pollen is a bit obscured by the variability within species, however species that are strong selfers or strong outcrossers have less variablity in mating systems and predictions of effect could be more realistic (see figure 1 from @whitehead2018)- IB this is a side problem for you, mention just once, and maybe in discusion. Now it has a lot of weigth.](). Moreover, other traits such as number of pollen grains per flower and number of ovules have been traditionally associated with the type of incompatibility system where species with higher pollen ovule ratios are predicted to be xenogamous and species with low pollen ovule ratios autogamous (REF) . Selfer species are known to have a reduction of herkogamy (REF) and less pollen production per ovule (REF) which can be interpretated as a reduction of pollen exported into the community [so, pollen and ovules are important by themselfs, or only because correlated? Unclear what your point is](). Other morphological traits, like stigma size can be determinant for the total pollen quatity that a stigma can receive and therefore related to do that pollen size would also play an important role. Example with pollen here.

[This last bit of the paragraph is still under development](Jose)

>>>>

@@

@@

Plant traits are crucial to understand heterospecific pollen effect, but here it is important to notice that these traits can be seen from a donor perpective or a recipient one. In first place, when the donor is considered pollen size, pollen allelopathy and the relatedness could be the two main predictors of effect. How pollen size and allepothaty are involved in heterospecific pollen effect remains unclear yet. Herrera xxxx observed that small pollen is more likely to clogg the stigma, here we argue that although this is possible, the quantities of heterospecific pollen in nature are generally low as mentioned above and bigger pollen is generally related with higher pollen tube growth rate, which add another layer of complexity to the understanding the overall importance of this trait. When the pollen recipeint is taking into account style length, stigmatic area, number of ovules and incompatibility system seems to be the main predictors of effect.

Example of flower morphology and the importance of stigmatic area Montgomery and rathcke 2012

Species that are self-incompatible have stronger barriers to heterospecific pollen than self-compatible species @ashman2013. Nonetheless, an effect of foreign pollen is a bit obscured by the variability within species, however species that are strong selfers or strong outcrossers have less variablity in mating systems and predictions of effect could be more realistic (see figure 1 from @whitehead2018).

End this with pollen quantity stuff!!!

Incompatibility system is another plant trait that appears to play an important role in foreign pollen effect. Species that are self incompatible have stronger barriers to heterospecific pollen than self-compatible species Ashman and Arceo-Gómez (2013). However, note the large variability in mating systems across populations Whitehead et al. (2018) . . . . The type of incompatibility, (i.e. whether sporophytic or gametophytic) is related to the location of pollen recognition; sporophytic incompatibility relates to signaling at the stigma surface while gametophytic occurs within the style Barrett (1988). This later acting pollen recognition mechanism is associated with greater negative effect than sporophytic recognistion Ashman and Arceo-Gómez (2013). if you introduce this,I eexpect you will use it in your analysis, remove (Nonetheless, an effect of foreign pollen is a bit obscured by the variability within species, however species that are strong selfers or strong outcrossers have less variablity in mating systems and predictions of effect could be more realistic (see figure 1 from
69 Whitehead et al. 2018)- IB this is a side problem for you, mention just once, and maybe in discusion.
70 Now it has a lot of weigth.)(). Moreover, other traits such as number of pollen grains per flower and
71 number of ovules have been traditionally associated with the type of incompatibility system where
72 species with higher pollen ovule ratios are predicted to be xenogamous and species with low pollen
73 ovule ratios autogamous (REF) . Selfer species are known to have a reduction of herkogamy (REF) and
74 less pollen production per ovule (REF) which can be interpretated as a reduction of pollen exported
75 into the community so, pollen and ovules are important by themselfs, or only because correlated? Unclear what your point is. Other morphological traits, like stigma size can be determinant for the total pollen quatity that a stigma can receive and therefore related to do that pollen size would also play an important role. Example with pollen here.

...

Moreover, other traits such as number of pollen grains per flower and number of ovules have been traditionally associated with the type of incompatibility system where species with higher pollen ovule ratios are predicted to be xenogamous and species with low pollen ovule ratios autogamous (REF) . Selfer species are known to have a reduction of herkogamy (REF) and less pollen production per ovule (REF) which can be interpretated as a reduction of pollen exported into the community [so, pollen and ovules are important by themselfs, or only because correlated? Unclear what your point is](). Other morphological traits, like stigma size can be determinant for the total pollen quatity that a stigma can receive and therefore related to do that pollen size would also play an important role. Example with pollen here.

### 2019-02-19 Manuscript_draft/Manuscript_draft.tex
some edits 3rd paragraph

\textbf{Paragraph 1} General idea to our concept

@@

\textbf{Paragraph 2} Introducing topic and knowledge gap

@@

\textbf{Paragraph 3} Expanding ideas with examples

Incompatibility system is another plant trait that appears to play an
important role in foreign pollen effect. Species that are self
incompatible have stronger barriers to heterospecific pollen than
self-compatible species Ashman and Arceo-Gómez (2013). However, note the
large variability in mating systems across populations Whitehead et al.
(2018) \href{}{\ldots{}}. The type of incompatibility, (i.e.~whether
sporophytic or gametophytic) is related to the location of pollen
recognition; sporophytic incompatibility relates to signaling at the
stigma surface while gametophytic occurs within the style Barrett
(1988). This later acting pollen recognition mechanism is associated
with greater negative effect than sporophytic recognistion Ashman and
Arceo-Gómez (2013). \href{}{if you introduce this,I eexpect you will use
it in your analysis, remove} (Nonetheless, an effect of foreign pollen
is a bit obscured by the variability within species, however species
that are strong selfers or strong outcrossers have less variablity in
mating systems and predictions of effect could be more realistic (see
figure 1 from Whitehead et al. 2018)- IB this is a side problem for you,
mention just once, and maybe in discusion. Now it has a lot of
weigth.)(). Moreover, other traits such as number of pollen grains per
flower and number of ovules have been traditionally associated with the
type of incompatibility system where species with higher pollen ovule
ratios are predicted to be xenogamous and species with low pollen ovule
ratios autogamous (REF) . Selfer species are known to have a reduction
of herkogamy (REF) and less pollen production per ovule (REF) which can
be interpretated as a reduction of pollen exported into the community
\href{}{so, pollen and ovules are important by themselfs, or only
because correlated? Unclear what your point is}. Other morphological
traits, like stigma size can be determinant for the total pollen quatity
that a stigma can receive and therefore related to do that pollen size
would also play an important role. Example with pollen here.

\href{Jose}{This last bit of the paragraph is still under development}

@@

\hypertarget{ref-barrett1988}{}
Barrett, S. C. 1988. The evolution, maintenance, and loss of
self-incompatibility systems. Plant reproductive ecology: patterns and
strategies:98--124.

>>>>

@@

@@

Plant traits are crucial to understand heterospecific pollen effect, but
here it is important to notice that these traits can be seen from a
donor perpective or a recipient one. In first place, when the donor is
considered pollen size, pollen allelopathy and the relatedness could be
the two main predictors of effect. How pollen size and allepothaty are
involved in heterospecific pollen effect remains unclear yet. Herrera
xxxx observed that small pollen is more likely to clogg the stigma, here
we argue that although this is possible, the quantities of
heterospecific pollen in nature are generally low as mentioned above and
bigger pollen is generally related with higher pollen tube growth rate,
which add another layer of complexity to the understanding the overall
importance of this trait. When the pollen recipeint is taking into
account style length, stigmatic area, number of ovules and
incompatibility system seems to be the main predictors of effect.

Example of flower morphology and the importance of stigmatic area
Montgomery and rathcke 2012

Species that are self-incompatible have stronger barriers to
heterospecific pollen than self-compatible species Ashman and
Arceo-Gómez (2013). Nonetheless, an effect of foreign pollen is a bit
obscured by the variability within species, however species that are
strong selfers or strong outcrossers have less variablity in mating
systems and predictions of effect could be more realistic (see figure 1
from Whitehead et al. (2018)).

End this with pollen quantity stuff!!!

Incompatibility system is another plant trait that appears to play an
important role in foreign pollen effect. Species that are self
incompatible have stronger barriers to heterospecific pollen than
self-compatible species Ashman and Arceo-Gómez (2013). However, note the
large variability in mating systems across populations Whitehead et al.
(2018) . . . . The type of incompatibility, (i.e.~whether sporophytic or
gametophytic) is related to the location of pollen recognition;
sporophytic incompatibility relates to signaling at the stigma surface
while gametophytic occurs within the style Barrett (1988). This later
acting pollen recognition mechanism is associated with greater negative
effect than sporophytic recognistion Ashman and Arceo-Gómez (2013). if
you introduce this,I eexpect you will use it in your analysis, remove
(Nonetheless, an effect of foreign pollen is a bit obscured by the
variability within species, however species that are strong selfers or
strong outcrossers have less variablity in mating systems and
predictions of effect could be more realistic (see figure 1 from 69
Whitehead et al. 2018)- IB this is a side problem for you, mention just
once, and maybe in discusion. 70 Now it has a lot of weigth.)().
Moreover, other traits such as number of pollen grains per flower and 71
number of ovules have been traditionally associated with the type of
incompatibility system where 72 species with higher pollen ovule ratios
are predicted to be xenogamous and species with low pollen 73 ovule
ratios autogamous (REF) . Selfer species are known to have a reduction
of herkogamy (REF) and 74 less pollen production per ovule (REF) which
can be interpretated as a reduction of pollen exported 75 into the
community so, pollen and ovules are important by themselfs, or only
because correlated? Unclear what your point is. Other morphological
traits, like stigma size can be determinant for the total pollen quatity
that a stigma can receive and therefore related to do that pollen size
would also play an important role. Example with pollen here.

\ldots{}

Moreover, other traits such as number of pollen grains per flower and
number of ovules have been traditionally associated with the type of
incompatibility system where species with higher pollen ovule ratios are
predicted to be xenogamous and species with low pollen ovule ratios
autogamous (REF) . Selfer species are known to have a reduction of
herkogamy (REF) and less pollen production per ovule (REF) which can be
interpretated as a reduction of pollen exported into the community
\href{}{so, pollen and ovules are important by themselfs, or only
because correlated? Unclear what your point is}. Other morphological
traits, like stigma size can be determinant for the total pollen quatity
that a stigma can receive and therefore related to do that pollen size
would also play an important role. Example with pollen here.

@@
